---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document:
    keep_md: yes
  pdf_document: default
---


## Loading and preprocessing the data

```{r}
library(plyr)
library(ggplot2)

temp <- tempfile()

download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", temp,"curl")

con <- unz(temp, "activity.csv")
activity <- read.table(con, header=T, sep=",")
unlink(temp)
clean <- activity[complete.cases(activity),]

dailyTotals <- ddply(clean,.(date),summarize,totalSteps=round(sum(steps),2))


IntervalTotals <- ddply(clean,.(interval),summarize,aveSteps=round(mean(steps),2))

emptyRecords  <- nrow(activity) - nrow(clean)

```

We have `r emptyRecords` records that are missing values.

## What is mean total number of steps taken per day?

```{r echo=FALSE}
plot1 <- qplot(data=dailyTotals,x=totalSteps,main="Distribution of Number of Steps",ylab="Number of days with these step counts",xlab="Steps take during a day",Nubinwidth=1000)

plot1
```

```{r echo=TRUE}
meanClean <- mean(dailyTotals$totalSteps)
medianClean <- median(dailyTotals$totalSteps)

```

We had a mean of `r format(meanClean, format="d", big.mark=',')` and medain of `r format(medianClean, format="d", big.mark=',')`.

## What is the average daily activity pattern?


```{r echo=FALSE}
plot2 <- qplot(data=IntervalTotals,type="l",x=interval,y=aveSteps,geom="line",ylab="Average Steps taking durning the Interval",xlab="15 minute interval during the Day") 
plot2

```

```{r echo=TRUE}
maxrow <-which.max(IntervalTotals[,2])
maxinterval <- IntervalTotals$interval[maxrow]
```

`r maxinterval` is the 5-minute interval, on average across all the days in the dataset, thate contains the maximum number of steps.


## Imputing missing values

From earlier we know that we have `r emptyRecords` records that are missing values.

We will transform the dataset by replacing NA with the median across intervals.

```{r echo=TRUE}

df <- activity

filled <- ddply(activity, 
      .(interval), 
      transform, 
      steps=ifelse(is.na(steps), median(steps, na.rm=TRUE), steps))
cleanFilled <- activity[complete.cases(filled),]
dailyFilledTotals <- ddply(filled,.(date),summarize,totalSteps=round(sum(steps),2))


emptyFilledRecords  <- nrow(activity) - nrow(cleanFilled)
meanFilledClean <- mean(dailyFilledTotals$totalSteps)
medianFilledClean <- median(dailyFilledTotals$totalSteps)


```



If we filled `r emptyFilledRecords` should be zero.

We had a mean of `r format(meanFilledClean, format="d", big.mark=',')` and medain of `r format(medianFilledClean, format="d", big.mark=',')` when we filled with the medians.

```{r echo=TRUE}
nrow(dailyFilledTotals)
nrow(dailyTotals)
```

Normally we would think that filling would increaase the mean, but because we had entirely empty days, when we filled for those it brought the average down.


```{r echo=FALSE}
plot1 <- qplot(data=dailyFilledTotals,x=totalSteps,main="Distribution of Number of Steps after NA",ylab="Number of days with these step counts",xlab="Steps take during a day",Nubinwidth=1000)

plot1
```

## Are there differences in activity patterns between weekdays and weekends?


```{r echo=TRUE}
clean$weekday <- !(weekdays(as.Date(clean$date)) %in% c('Saturday','Sunday')) 

IntervalAbbeyTotals <- ddply(clean,.(interval,weekday),summarize,aveSteps=round(mean(steps),2))
```


```{r echo=FALSE}
plotPanel <- qplot(data=IntervalAbbeyTotals,x=interval,y=aveSteps,geom="line",ylab="Number of Steps",xlab="Interval") + facet_wrap(  ~  weekday, ncol=1)

plotPanel

```

